### 2022-02-13 TP1/tex/plots/df_bis_ucmp_reg.tex

% Created by tikzDevice version 0.12.3.1 on 2022-02-12 22:24:19
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (505.89,216.81);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (505.89,216.81);
\end{scope}
\begin{scope}
\path[clip] ( 44.91, 30.69) rectangle (500.39,178.94);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] ( 44.91, 30.69) rectangle (500.39,178.94);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.91, 57.36) --
	(500.39, 57.36);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.91, 95.30) --
	(500.39, 95.30);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.91,133.24) --
	(500.39,133.24);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.91,171.18) --
	(500.39,171.18);

\path[draw=drawColor,line width= 0.3pt,line join=round] (133.28, 30.69) --
	(133.28,178.94);

\path[draw=drawColor,line width= 0.3pt,line join=round] (269.05, 30.69) --
	(269.05,178.94);

\path[draw=drawColor,line width= 0.3pt,line join=round] (404.83, 30.69) --
	(404.83,178.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.91, 38.39) --
	(500.39, 38.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.91, 76.33) --
	(500.39, 76.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.91,114.27) --
	(500.39,114.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.91,152.21) --
	(500.39,152.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 65.39, 30.69) --
	( 65.39,178.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (201.16, 30.69) --
	(201.16,178.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.94, 30.69) --
	(336.94,178.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (472.72, 30.69) --
	(472.72,178.94);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 65.61, 38.44) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 67.30, 38.88) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 69.35, 39.46) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 71.58, 40.11) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 73.94, 40.81) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 76.39, 41.53) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 78.91, 42.30) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 81.49, 43.15) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 84.13, 43.90) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 86.82, 44.74) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 89.55, 45.53) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 92.31, 46.37) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 95.12, 47.25) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 97.95, 48.16) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (100.81, 49.00) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (103.70, 50.03) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (106.62, 50.74) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (109.56, 51.85) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (112.53, 52.65) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (115.51, 53.61) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (118.52, 54.67) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (121.54, 55.66) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (124.59, 56.67) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (127.65, 57.76) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (130.72, 58.46) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (133.82, 59.45) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (136.93, 60.19) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (140.05, 61.31) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (143.19, 62.41) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (146.34, 63.43) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (149.51, 64.64) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (152.69, 65.39) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (155.88, 66.45) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (159.08, 67.23) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (162.30, 68.38) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (165.53, 69.41) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (168.76, 70.74) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (172.01, 71.73) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (175.27, 72.71) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (178.54, 73.95) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (181.82, 74.85) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (185.11, 75.88) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (188.41, 77.21) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (191.72, 78.07) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (195.04, 79.05) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (198.36, 80.13) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (201.70, 81.20) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (205.04, 82.59) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (208.39, 82.89) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (211.75, 84.43) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (215.12, 85.59) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (218.49, 86.78) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (221.88, 87.76) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (225.27, 88.81) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (228.67, 90.32) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (232.07, 91.39) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (235.48, 92.02) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (238.90, 93.41) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (242.33, 94.11) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (245.76, 95.29) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (249.20, 96.41) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (252.65, 97.93) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (256.10, 99.25) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (259.56,100.19) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (263.02,100.87) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (266.49,102.39) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (269.97,103.42) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (273.45,104.18) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (276.94,105.99) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (280.43,107.20) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (283.93,107.79) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (287.44,108.99) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (290.95,109.74) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (294.47,111.44) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (297.99,112.86) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (301.52,113.29) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (305.05,115.18) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (308.59,115.56) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (312.13,117.21) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (315.68,117.95) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (319.23,119.36) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (322.79,120.15) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (326.35,121.82) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (329.92,122.92) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (333.49,124.78) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (337.06,125.57) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (340.64,126.13) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (344.23,128.70) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (347.82,128.57) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (351.41,129.96) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (355.01,131.03) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (358.62,131.30) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (362.22,133.74) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (365.83,134.82) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (369.45,135.12) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (373.07,137.01) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (376.69,138.32) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (380.32,139.51) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (383.95,140.80) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (387.59,141.42) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (391.23,143.08) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (394.88,144.79) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (398.52,145.43) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (402.17,146.78) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (405.83,148.52) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (409.49,149.52) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (413.15,148.98) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (416.82,151.23) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (420.49,152.32) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (424.16,152.94) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (427.84,154.94) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (431.52,157.06) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (435.21,156.80) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (438.90,158.33) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (442.59,161.44) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (446.28,160.66) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (449.98,162.73) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (453.68,163.43) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (457.39,165.82) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (461.10,166.73) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (464.81,166.73) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (468.52,168.30) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (472.24,169.41) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (475.96,170.50) circle (  0.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (479.69,172.20) circle (  0.62);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 65.61, 37.42) --
	( 67.30, 37.97) --
	( 69.35, 38.64) --
	( 71.58, 39.36) --
	( 73.94, 40.12) --
	( 76.39, 40.92) --
	( 78.91, 41.73) --
	( 81.49, 42.57) --
	( 84.13, 43.43) --
	( 86.82, 44.30) --
	( 89.55, 45.18) --
	( 92.31, 46.08) --
	( 95.12, 46.99) --
	( 97.95, 47.91) --
	(100.81, 48.83) --
	(103.70, 49.77) --
	(106.62, 50.72) --
	(109.56, 51.67) --
	(112.53, 52.63) --
	(115.51, 53.60) --
	(118.52, 54.57) --
	(121.54, 55.55) --
	(124.59, 56.54) --
	(127.65, 57.53) --
	(130.72, 58.53) --
	(133.82, 59.53) --
	(136.93, 60.54) --
	(140.05, 61.55) --
	(143.19, 62.57) --
	(146.34, 63.59) --
	(149.51, 64.62) --
	(152.69, 65.65) --
	(155.88, 66.68) --
	(159.08, 67.72) --
	(162.30, 68.76) --
	(165.53, 69.81) --
	(168.76, 70.86) --
	(172.01, 71.91) --
	(175.27, 72.97) --
	(178.54, 74.03) --
	(181.82, 75.09) --
	(185.11, 76.16) --
	(188.41, 77.23) --
	(191.72, 78.30) --
	(195.04, 79.37) --
	(198.36, 80.45) --
	(201.70, 81.53) --
	(205.04, 82.62) --
	(208.39, 83.70) --
	(211.75, 84.79) --
	(215.12, 85.88) --
	(218.49, 86.98) --
	(221.88, 88.07) --
	(225.27, 89.17) --
	(228.67, 90.27) --
	(232.07, 91.38) --
	(235.48, 92.48) --
	(238.90, 93.59) --
	(242.33, 94.70) --
	(245.76, 95.81) --
	(249.20, 96.93) --
	(252.65, 98.05) --
	(256.10, 99.17) --
	(259.56,100.29) --
	(263.02,101.41) --
	(266.49,102.53) --
	(269.97,103.66) --
	(273.45,104.79) --
	(276.94,105.92) --
	(280.43,107.05) --
	(283.93,108.19) --
	(287.44,109.32) --
	(290.95,110.46) --
	(294.47,111.60) --
	(297.99,112.74) --
	(301.52,113.89) --
	(305.05,115.03) --
	(308.59,116.18) --
	(312.13,117.33) --
	(315.68,118.48) --
	(319.23,119.63) --
	(322.79,120.78) --
	(326.35,121.94) --
	(329.92,123.09) --
	(333.49,124.25) --
	(337.06,125.41) --
	(340.64,126.57) --
	(344.23,127.73) --
	(347.82,128.89) --
	(351.41,130.06) --
	(355.01,131.23) --
	(358.62,132.39) --
	(362.22,133.56) --
	(365.83,134.73) --
	(369.45,135.91) --
	(373.07,137.08) --
	(376.69,138.25) --
	(380.32,139.43) --
	(383.95,140.61) --
	(387.59,141.79) --
	(391.23,142.97) --
	(394.88,144.15) --
	(398.52,145.33) --
	(402.17,146.51) --
	(405.83,147.70) --
	(409.49,148.88) --
	(413.15,150.07) --
	(416.82,151.26) --
	(420.49,152.45) --
	(424.16,153.64) --
	(427.84,154.83) --
	(431.52,156.02) --
	(435.21,157.22) --
	(438.90,158.41) --
	(442.59,159.61) --
	(446.28,160.81) --
	(449.98,162.01) --
	(453.68,163.21) --
	(457.39,164.41) --
	(461.10,165.61) --
	(464.81,166.81) --
	(468.52,168.02) --
	(472.24,169.22) --
	(475.96,170.43) --
	(479.69,171.64);
\definecolor{drawColor}{RGB}{255,0,0}

\path[draw=drawColor,draw opacity=0.30,line width= 0.6pt,line join=round] ( 65.61, 38.44) --
	( 67.30, 38.88) --
	( 69.35, 39.46) --
	( 71.58, 40.11) --
	( 73.94, 40.81) --
	( 76.39, 41.53) --
	( 78.91, 42.30) --
	( 81.49, 43.15) --
	( 84.13, 43.90) --
	( 86.82, 44.74) --
	( 89.55, 45.53) --
	( 92.31, 46.37) --
	( 95.12, 47.25) --
	( 97.95, 48.16) --
	(100.81, 49.00) --
	(103.70, 50.03) --
	(106.62, 50.74) --
	(109.56, 51.85) --
	(112.53, 52.65) --
	(115.51, 53.61) --
	(118.52, 54.67) --
	(121.54, 55.66) --
	(124.59, 56.67) --
	(127.65, 57.76) --
	(130.72, 58.46) --
	(133.82, 59.45) --
	(136.93, 60.19) --
	(140.05, 61.31) --
	(143.19, 62.41) --
	(146.34, 63.43) --
	(149.51, 64.64) --
	(152.69, 65.39) --
	(155.88, 66.45) --
	(159.08, 67.23) --
	(162.30, 68.38) --
	(165.53, 69.41) --
	(168.76, 70.74) --
	(172.01, 71.73) --
	(175.27, 72.71) --
	(178.54, 73.95) --
	(181.82, 74.85) --
	(185.11, 75.88) --
	(188.41, 77.21) --
	(191.72, 78.07) --
	(195.04, 79.05) --
	(198.36, 80.13) --
	(201.70, 81.20) --
	(205.04, 82.59) --
	(208.39, 82.89) --
	(211.75, 84.43) --
	(215.12, 85.59) --
	(218.49, 86.78) --
	(221.88, 87.76) --
	(225.27, 88.81) --
	(228.67, 90.32) --
	(232.07, 91.39) --
	(235.48, 92.02) --
	(238.90, 93.41) --
	(242.33, 94.11) --
	(245.76, 95.29) --
	(249.20, 96.41) --
	(252.65, 97.93) --
	(256.10, 99.25) --
	(259.56,100.19) --
	(263.02,100.87) --
	(266.49,102.39) --
	(269.97,103.42) --
	(273.45,104.18) --
	(276.94,105.99) --
	(280.43,107.20) --
	(283.93,107.79) --
	(287.44,108.99) --
	(290.95,109.74) --
	(294.47,111.44) --
	(297.99,112.86) --
	(301.52,113.29) --
	(305.05,115.18) --
	(308.59,115.56) --
	(312.13,117.21) --
	(315.68,117.95) --
	(319.23,119.36) --
	(322.79,120.15) --
	(326.35,121.82) --
	(329.92,122.92) --
	(333.49,124.78) --
	(337.06,125.57) --
	(340.64,126.13) --
	(344.23,128.70) --
	(347.82,128.57) --
	(351.41,129.96) --
	(355.01,131.03) --
	(358.62,131.30) --
	(362.22,133.74) --
	(365.83,134.82) --
	(369.45,135.12) --
	(373.07,137.01) --
	(376.69,138.32) --
	(380.32,139.51) --
	(383.95,140.80) --
	(387.59,141.42) --
	(391.23,143.08) --
	(394.88,144.79) --
	(398.52,145.43) --
	(402.17,146.78) --
	(405.83,148.52) --
	(409.49,149.52) --
	(413.15,148.98) --
	(416.82,151.23) --
	(420.49,152.32) --
	(424.16,152.94) --
	(427.84,154.94) --
	(431.52,157.06) --
	(435.21,156.80) --
	(438.90,158.33) --
	(442.59,161.44) --
	(446.28,160.66) --
	(449.98,162.73) --
	(453.68,163.43) --
	(457.39,165.82) --
	(461.10,166.73) --
	(464.81,166.73) --
	(468.52,168.30) --
	(472.24,169.41) --
	(475.96,170.50) --
	(479.69,172.20);
\definecolor{drawColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 65.61, 37.42) -- ( 65.61, 38.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.30, 37.97) -- ( 67.30, 38.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 69.35, 38.64) -- ( 69.35, 39.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 71.58, 39.36) -- ( 71.58, 40.11);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 73.94, 40.12) -- ( 73.94, 40.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 76.39, 40.92) -- ( 76.39, 41.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.91, 41.73) -- ( 78.91, 42.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 81.49, 42.57) -- ( 81.49, 43.15);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 84.13, 43.43) -- ( 84.13, 43.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 86.82, 44.30) -- ( 86.82, 44.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 89.55, 45.18) -- ( 89.55, 45.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 92.31, 46.08) -- ( 92.31, 46.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 95.12, 46.99) -- ( 95.12, 47.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 97.95, 47.91) -- ( 97.95, 48.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (100.81, 48.83) -- (100.81, 49.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (103.70, 49.77) -- (103.70, 50.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (106.62, 50.72) -- (106.62, 50.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (109.56, 51.67) -- (109.56, 51.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.53, 52.63) -- (112.53, 52.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (115.51, 53.60) -- (115.51, 53.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (118.52, 54.57) -- (118.52, 54.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (121.54, 55.55) -- (121.54, 55.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (124.59, 56.54) -- (124.59, 56.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (127.65, 57.53) -- (127.65, 57.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.72, 58.53) -- (130.72, 58.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] (133.82, 59.53) -- (133.82, 59.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (136.93, 60.54) -- (136.93, 60.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (140.05, 61.55) -- (140.05, 61.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.19, 62.57) -- (143.19, 62.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (146.34, 63.59) -- (146.34, 63.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (149.51, 64.62) -- (149.51, 64.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (152.69, 65.65) -- (152.69, 65.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (155.88, 66.68) -- (155.88, 66.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.08, 67.72) -- (159.08, 67.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.30, 68.76) -- (162.30, 68.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.53, 69.81) -- (165.53, 69.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (168.76, 70.86) -- (168.76, 70.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (172.01, 71.91) -- (172.01, 71.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (175.27, 72.97) -- (175.27, 72.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (178.54, 74.03) -- (178.54, 73.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (181.82, 75.09) -- (181.82, 74.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (185.11, 76.16) -- (185.11, 75.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (188.41, 77.23) -- (188.41, 77.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (191.72, 78.30) -- (191.72, 78.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.04, 79.37) -- (195.04, 79.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.36, 80.45) -- (198.36, 80.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (201.70, 81.53) -- (201.70, 81.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.04, 82.62) -- (205.04, 82.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (208.39, 83.70) -- (208.39, 82.89);

\path[draw=drawColor,line width= 0.6pt,line join=round] (211.75, 84.79) -- (211.75, 84.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (215.12, 85.88) -- (215.12, 85.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (218.49, 86.98) -- (218.49, 86.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (221.88, 88.07) -- (221.88, 87.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (225.27, 89.17) -- (225.27, 88.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.67, 90.27) -- (228.67, 90.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (232.07, 91.38) -- (232.07, 91.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (235.48, 92.48) -- (235.48, 92.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (238.90, 93.59) -- (238.90, 93.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (242.33, 94.70) -- (242.33, 94.11);

\path[draw=drawColor,line width= 0.6pt,line join=round] (245.76, 95.81) -- (245.76, 95.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (249.20, 96.93) -- (249.20, 96.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (252.65, 98.05) -- (252.65, 97.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (256.10, 99.17) -- (256.10, 99.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (259.56,100.29) -- (259.56,100.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (263.02,101.41) -- (263.02,100.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] (266.49,102.53) -- (266.49,102.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (269.97,103.66) -- (269.97,103.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (273.45,104.79) -- (273.45,104.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (276.94,105.92) -- (276.94,105.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (280.43,107.05) -- (280.43,107.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (283.93,108.19) -- (283.93,107.79);

\path[draw=drawColor,line width= 0.6pt,line join=round] (287.44,109.32) -- (287.44,108.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (290.95,110.46) -- (290.95,109.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.47,111.60) -- (294.47,111.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (297.99,112.74) -- (297.99,112.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (301.52,113.89) -- (301.52,113.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (305.05,115.03) -- (305.05,115.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (308.59,116.18) -- (308.59,115.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.13,117.33) -- (312.13,117.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (315.68,118.48) -- (315.68,117.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (319.23,119.63) -- (319.23,119.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (322.79,120.78) -- (322.79,120.15);

\path[draw=drawColor,line width= 0.6pt,line join=round] (326.35,121.94) -- (326.35,121.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (329.92,123.09) -- (329.92,122.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.49,124.25) -- (333.49,124.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.06,125.41) -- (337.06,125.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (340.64,126.57) -- (340.64,126.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.23,127.73) -- (344.23,128.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (347.82,128.89) -- (347.82,128.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (351.41,130.06) -- (351.41,129.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (355.01,131.23) -- (355.01,131.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (358.62,132.39) -- (358.62,131.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (362.22,133.56) -- (362.22,133.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (365.83,134.73) -- (365.83,134.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.45,135.91) -- (369.45,135.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (373.07,137.08) -- (373.07,137.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (376.69,138.25) -- (376.69,138.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (380.32,139.43) -- (380.32,139.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (383.95,140.61) -- (383.95,140.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (387.59,141.79) -- (387.59,141.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (391.23,142.97) -- (391.23,143.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (394.88,144.15) -- (394.88,144.79);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.52,145.33) -- (398.52,145.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (402.17,146.51) -- (402.17,146.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.83,147.70) -- (405.83,148.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (409.49,148.88) -- (409.49,149.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (413.15,150.07) -- (413.15,148.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.82,151.26) -- (416.82,151.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.49,152.45) -- (420.49,152.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (424.16,153.64) -- (424.16,152.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (427.84,154.83) -- (427.84,154.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (431.52,156.02) -- (431.52,157.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (435.21,157.22) -- (435.21,156.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (438.90,158.41) -- (438.90,158.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] (442.59,159.61) -- (442.59,161.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (446.28,160.81) -- (446.28,160.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (449.98,162.01) -- (449.98,162.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (453.68,163.21) -- (453.68,163.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (457.39,164.41) -- (457.39,165.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (461.10,165.61) -- (461.10,166.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.81,166.81) -- (464.81,166.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.52,168.02) -- (468.52,168.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (472.24,169.22) -- (472.24,169.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (475.96,170.43) -- (475.96,170.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (479.69,171.64) -- (479.69,172.20);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 39.96, 35.36) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 39.96, 73.30) {20000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 39.96,111.24) {40000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 39.96,149.18) {60000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 42.16, 38.39) --
	( 44.91, 38.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 42.16, 76.33) --
	( 44.91, 76.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 42.16,114.27) --
	( 44.91,114.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 42.16,152.21) --
	( 44.91,152.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 65.39, 27.94) --
	( 65.39, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (201.16, 27.94) --
	(201.16, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.94, 27.94) --
	(336.94, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (472.72, 27.94) --
	(472.72, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 65.39, 19.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (201.16, 19.68) {20000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (336.94, 19.68) {40000};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (472.72, 19.68) {60000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (272.65,  7.64) {$n \log n$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 44.91,186.58) {$\texttt{nombreComparaisons}(n) \approx 1.16(n\log n) - 549.56$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 44.91,202.22) {Moyenne des comparaisons};
\end{scope}
\end{tikzpicture}